---
title: "SMSCG action plots"
author: "Rosie"
date: "7/15/2020"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(DataCombine)
library(lubridate)
library(visreg)
library(ggthemes)
library(nlme)

```

## Data loading
Firt I loaded the data that you sent me earlier, and set up a data frame with just the years we are interested in

```{r loading}
#load the data from the sondes
load("~/salinity control gates/SMSCG/actiondata.RData")
load("~/salinity control gates/SMSCG/waterquality.RData")
load("~/salinity control gates/SMSCG/actiondata-october.RData")

#create a data frame of year types
yrtyps = data.frame(an= c(2005, 2006, 2017), bn= c(2002, 2009, 2016)) %>% 
  gather(key = "yrtyp", value = "Year")
```

Now I did a little manipulation on the action data and the historical data to put it in the right format. Then I summarized by month to calculate the standard error

```{r}
#put all the 2018 data together
action.daily2 = rbind(action.daily, action.daily.october)
action.daily2$Month = month(action.daily2$Datetime)

action.monthly = group_by(action.daily2, Station, Analyte, Month) %>% 
  summarize(Meanx = mean(Mean), sd = sd(Mean)) %>%
  mutate(yrtyp = "action") %>%
  filter(Month != 11)

```

Then I did the same thing for the historic data and bound the two data sets together

```{r}

historical.daily = mutate(historical.daily, Year = year(Datetime),
                          Month = month(Datetime))

#add the year types
historical.daily2 = merge(historical.daily, yrtyps)

#Calculate the monthly means and standard deviations
historical.monthly =  group_by(filter(historical.daily2, !is.na(Mean), Year != 2018 & Year !=2019), 
                               Station, Analyte, Month, yrtyp) %>% 
  summarize(Meanx = mean(Mean, na.rm = T), sd = sd(Mean, na.rm=T)) %>%
  filter(Month %in% c(7,8,9,10))


alldata = rbind(historical.monthly, action.monthly)

# filter out just the stations we are interested in
alldata = filter(alldata, Station == "(C-2B)  Collinsville B" | Station == "(S-64)  National Steel" |
                   Station =="(S-54)  Hunter Cut") %>%
  mutate(Station = factor(Station, levels = c("(S-54)  Hunter Cut", "(S-64)  National Steel" ,"(C-2B)  Collinsville B"),
                         labels = c("West Marsh", "East Marsh", "River")),
         Month = factor(Month, levels = c(7, 8, 9, 10),
                         labels = c("Jul", "Aug", "Sep", "Oct")),
         Analyte = factor(Analyte, levels = c("Salinity","Temperature","Fluorescence","Turbidity")),
         yrtyp = factor(yrtyp, levels = c("bn","an","action")))



```



## Plots


```{r}

p5 = ggplot(data = alldata, aes(x=Month, y = Meanx, fill = yrtyp))
p5.1 = p5 + geom_bar(stat = "identity", position = "dodge", color = "black") + 
  geom_errorbar(aes(ymin = Meanx + sd, ymax = Meanx - sd), group = "yrtyp", 
                position = position_dodge(.9), width = .5)+
  facet_grid(Analyte~Station, scales = "free_y", space = "free_x")+
  scale_fill_manual(values = c(
    "bn" = "#dfc27d",
    "an" = "#92c5de",
    "action" = "#e31a1c"),
    labels = c("dry summers",
                                 "wet summers","2018"), name = NULL)+
  ylab("Turbidity (NTU)      Chlorophyll (ug/L)    Temperature (C)     Salinity (PSU)")+
  theme_bw()+ theme(legend.position="bottom", strip.background.y = element_blank(),
                     strip.text.y = element_blank(), axis.title.x = element_blank())

p5.1

```

