---
title: "FMWT_SMSCG"
author: "Rosemary Hartman"
date: "7/12/2019"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}

library(tidyverse)
library(readxl)
library(lme4)
library(lmerTest)
library(lubridate)
library(visreg)
library(pscl)
library(MASS)
source("data manip FMWT.R")
```

In the Delta Smelt Resiliancy strategy, there is the idea we can change the operation of the Suisun Marsh Salinity control gates to improve habitat in the marsh for Delta Smelt. I was curious if there were any trends between historical gate operations and presence of Delta Smelt in the Marsh. Therefore, I decided to compare catch of Delta Smelt in the marsh during the fall (when the gates are most frequently operated) with gate operations.

First I did some data manipulation to get the gate operation data lined up with the Delta Smelt Catch from the Fall Midwater Trawl, and I calculated CPUE.

FMWT data is avaialable here: ftp://ftp.wildlife.ca.gov/TownetFallMidwaterTrawl/FMWT%20Data/
Michael Koohafman gave me the gate operation data.

First for some quick exploritory plots of the data. Look at when the gates are usually operated and what the fish catch was like when they are or are not operated.

```{r opsplots, echo = TRUE}

#when are the gates operated?
ggplot(op.daily, aes(x=julian, fill = Operating2)) + geom_bar(stat = "Count")

#look at it a different way
 ggplot(op.daily, aes(x=julian, fill = Operating2)) + geom_bar(stat = "Count", position = "fill") +
    facet_wrap(.~Year, scales = "free") + scale_y_discrete(breaks = NULL)
```
 How does gate operation relate to water year type?
   I'll make a plot and run a basic model to see whether gates are operated more frequently in dry years. I've binned the five year types into "wet" and "dry"
```{r opsplots2, echo = T}
ops = filter(merge(op.daily, yrtyp), Year >1988, Year <2019)
ops$Operating2 = as.factor(ops$Operating2)

#summarize number of days per year the gates were operated, grouping by year type
opsum = group_by(ops, Year, YT2) %>% summarize(ops = length(Operating2[which(Operating2 == "Operating")]))
opsmean = group_by(opsum, YT2) %>% summarize(mean = mean(ops), se = sd(ops)/length(ops), sd = sd(ops))

#plot it
ggplot(opsmean, aes(x=YT2, y= mean, fill = YT2)) + geom_bar(stat = "identity") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se)) + ylab("days of operation") +
  scale_x_discrete(name = "Year type", labels = c("Dry", "Wet"))

#It looks like they are operated more often in dry years, but I'll stats it to check.
opmod2 = glm(ops~ YT2, data = opsum)
summary(opmod2)
visreg(opmod2)
#Yup. Definitely. 

```
Now I"ll make some basic plots of fish catch. I'll filter it to just the time period where we have gate data (1988-2011) and just the fall, since most of the FMWT data is Sep-Dec. I didn't plot 2012-2018, because the catch was too low
```{r plotsfish, echo = T}

ggplot(FMWT_DSmg2, aes(x = Operating2, y = log(CPUE+1))) + geom_boxplot()

#seperate by month
ggplot(FMWT_DSmg2, aes(x = Operating2, y = log(catch+1))) + geom_boxplot() + facet_wrap(~month(Date))

#Look at smelt catch in Montezuma Slough and the FMWT index overall - do they match up?
ggplot(tots, aes(x=Year, y= catch)) + geom_bar(stat = "identity")
ggplot(index, aes(x=Year, y= Total)) + geom_bar(stat = "identity")


```

Let's run some models to see whether there are statistically more smelt when the gates are operating. Other things are probably involved too, such as year, water year type, salinity, station, day of the year, etc. I'll run several models and rank them with AICc to see which is best.

It took me a long time to figure out what type of model to run. Delta smelt catch data is "count data", so theoretically it should follow a poisson distribution. However, my preliminary analysis showed it is highly overdisperssed and has WAY more zeros than a normal Poisson distribution. Therefore, after much discussion, research, statistics textbooks, and false starts, I settled on a zero-inflated negative binomial model.

I added the volume sampled as an offset, since I want count data rather than CPUE, but I still want to encorporate the volume's potential effect on catch. 

Now for some models. Some notes:
    - I have "scaled" all the predictor variables so that they have a mean of zero and a standard deviation of 1. This lets me compare between variables more easily.
    - I subset the data so that it is just dry years (below normal, dry, or critical) prior to 2011
    - I binned the gate operation into "operating" or "not", I did not make any distinction bewtween full and partial operation.
    - EC is the conductivity measured by CDFW's boat, not the nearest WQ station.
    - SacX2 is from Hutton et al. 
    - If no volume of sample was available, I replaced it with the average.
```{r models, echo = T}

#First I'll put in all the variables for a global model
dsznb2 = zeroinfl(catch~ #catch of Delta Smelt
                    Station+ #Station on Montezuma Slough
                    Operating2+ #whether the gates are operating
                    julianscaled + #Day of the Year (scaled)
                    Indexscaled + #FMWT index (scaled)
                    SacX2 + #sacramento river X2
                    ECscaled + #Top electrical conductivity (scaled)
                    offset(Vol), #volume added as an offset
                  dist = "negbin", data = FMWT_DSmg4a, na.action = "na.fail")

#assess all possible models
dreznb2 = dredge(dsznb2)
head(dreznb2)

#this was the best model
dsznb2best = zeroinfl(catch~ Operating2+julianscaled + 
                        ECscaled + Indexscaled+ offset(Vol),
                      dist = "negbin", data = FMWT_DSmg4a, na.action = "na.fail")
summary(dsznb2best)

#I can make some conditional plots to see the effect of each variable by itself.
visreg(dsznb2best)

```
So that model looks OK, but we should check the residuals to make sure there is no pattern. According to Zuur et al. 2009, I should plot the pearson residuals versus the fitted values and versus the predictor variables and see no pattern.
```{r plotsresid, echo = T}

tests = data.frame(DSresid = residuals(dsznb2best, type = "pearson"), 
                   DSfit = dsznb2best$fitted.values,
                   ops = dsznb2best$model["Operating2"],
                   EC = dsznb2best$model["ECscaled"],
                   Ind = dsznb2best$model["Indexscaled"],
                   day = dsznb2best$model["julianscaled"],
                   catch = FMWT_DSmg4a$catch)


ggplot(data = filter(tests, catch !=0), aes(x=DSresid, y = DSfit)) + geom_point()
ggplot(data = tests, aes(x=DSresid, y = ECscaled)) + geom_point()
ggplot(data = tests, aes(x=DSresid, y = Operating2)) + geom_point()
ggplot(data = tests, aes(x=DSresid, y = julianscaled)) + geom_point()
ggplot(data = tests, aes(x=DSresid, y = Indexscaled)) + geom_point()

```
If we've been operating the gates whenever it gets salty in the marsh, It would be ideal to see what would have happened to smelt catch in a dry year in the marsh where we didn't operate the gates.

Fortunately, we have dtata from FMWT before the gates went into operation.
```{r premodels, echo = T}

###########################################################################################
#can we compare pre-gate data to post-gate data?

#first make a new variable for pre-gate versus post-gate
FMWT_DSmg4ax$prepost = NA
FMWT_DSmg4ax$prepost[which(FMWT_DSmg4ax$Year <1988)] = "pre"
FMWT_DSmg4ax$prepost[which(is.na(FMWT_DSmg4ax$prepost))] = "post"

#Take out the couple of places were we didn't have X2 values
FMWT_DSmg4ax = filter(FMWT_DSmg4ax,!is.na(SacX2))

#quick plot of smelt abundacne pre versus post
ggplot(FMWT_DSmg4ax, aes(x=prepost, y = log(catch+1))) + geom_boxplot()
ggplot(FMWT_DSmg4ax, aes(x=prepost, y = log(CPUE+1))) + geom_boxplot()


#run a model with all the possible variables
dsznb3 = zeroinfl(catch~Station + prepost + julianscaled + 
                    ECscaled +SacX2 + Indexscaled + offset(Vol), dist = "negbin", 
                  data = FMWT_DSmg4ax,na.action = "na.fail")

#What does this one look like?
summary(dsznb3)

#Go through all the possible models
drepp = dredge(dsznb3)
head(drepp)
#It liked the full model, except for the volume offset. 

visreg(dsznb3,  gg = T)

#Try it again without the volume offset
dsznb4 = zeroinfl(catch~Station + prepost + julianscaled + 
                    ECscaled +SacX2 + Indexscaled, dist = "negbin", 
                  data = FMWT_DSmg4ax,na.action = "na.fail")
summary(dsznb4)
visreg(dsznb4)

```
OK, That's wierd the pre/post coefficient reversed.I think we'll stick with the offset. Anyway, It doesn't look like this is a particularly useful avenue for evaluating the gates, since Delta Smelt were a lot more common everywhere in the 70s and early 80s than they were in the 90s and 2000s
